\documentclass[english]{article}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Finding the best meeting point}
\author{johmathe@google.com}
\maketitle

\section{Problem}

Given $N$ people in a manhattan-like city, find the best meeting
point for all these people (they are not trying to score fitbit points,
they just want to optimize their trajectory).


\section{A failed attempt}

Let's take the problem in one dimension as a first simplification
(it is actually a separable problem). First idea: take the mean of
all the coordinates. This is not the optimal solution, and there is
a simple counter example for this: imagine 9 people are at coordinate
0 and one other person is at 10. This means that people should meet
at:

\[
\bar{x}=\frac{1}{N}\sum_{i=1}^{N}x_{i}=1.
\]


The total distance that people need to travel is then:

\[
d=\sum_{i=1}^{N}|x_{i}-\bar{x}|=18
\]


An obvious counter example is $\tilde{x}=0$. In that case, only the
last person needs to travel to this meeting point, and the distance
becomes:

\[
\tilde{d}=\sum_{i=1}^{N}|x-\tilde{x}|=10
\]


Since $d>\tilde{d}$ we can say that the mean doesn't give the best
meeting point.


\section{Proper demonstration}

We build an objective (cost) function that returns the total distance
travelled by people, with $\bar{x}$ and $\bar{y}$ our coordinates.
Our goal will be to minimize it.

\[
\mbox{minimize }f(\bar{x},\bar{y})=\sum_{i=1}^{N}|x_{i}-x|+\sum_{i=1}^{N}|y_{i}-y|
\]


We see here that the function is separable. So we can minimize independently
accross $\bar{x}$ and $\bar{y}$. 

This function is (non-strictly) convex. In order to mininmize it,
we can find the subgradient set of this function, and find for which
value(s) of $\bar{x}$ and $\bar{y}$ it is equal to $0$ among the
two axes - we use the subgradient since our function is not continous
and doesn't have a derivative defined on the non-continous points%
\footnote{This is just a generalization of the usual method of finding the global
minimum of a continuous and strictly convex function (take the derivative and equate to 0)%
}:

\[
G(\bar{x})=\sum_{i=1}^{N}\mbox{sign}(\mbox{\ensuremath{\bar{x}}}-x_{i})=0
\]


The solution for $\bar{x}$ and $\bar{y}$ is the median in this case
(drawing a graph helps seeing why).

\[
\bar{x}=\mu_{\frac{1}{2}}(x)
\]


\[
\bar{y}=\mu_{\frac{1}{2}}(y)
\]



\section{Solving the problem and complexity}

Now that we know that the problem is just a median finding problem,
we can take a simple approach which consists in sorting the data points
with a well known alogrithm, and take the element from the middle
if we have an odd number of elements, or average the two elements
in the middle of the array if we have an even number of elements%
\footnote{See definition of the median if rusty :)%
}. This approach is $O(N\log N)$in average (the best theoeretical
bound for comparaison based algorithms is $O(N\log N)$). We could
also do better with a radix sort if all our coordinates are integers
and get down to $O(N)$. There are also some median finding algorithms
out there that have a complexity of $O(N)$.
\end{document}

